---
title: "V1 DSP vs TOP Analysis"
author: "Lorenzo Tubertini"
date: "17 April 2018"
output:
  github_document:
    dev: jpeg
    toc: true
    toc_depth: 3

---

## R code to prepare the V1 DSP vs TOP Analysis

### Libraries load, MongoDB data import and consolidation
CSV files from the original case can be found [here](https://github.com/ltubertini/DSP_TOP/tree/master/case)

```{r data import, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(vcd)

setwd('C:/Users/ltubertini/Downloads/case')
a <- (read.csv2('case.csv', sep=";", na.strings = "-",dec=',', skip=1,
               colClasses=c("character","character","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric")))
b <- (read.csv2('case2.csv', sep=";", na.strings = "-",dec=',', skip=1,
               colClasses=c("character","character","numeric")))

input <- merge(a, b, by=c("ad_id","dealer_id"), all=T)
input <- input %>%
  group_by(ad_id,dealer_id) %>%
  summarise_all(sum)
```

### Ad type bucketing, subsetting
Ads are bucketed as "DSP", "TOP"" or both ("multi") based on the available flags for readibility.
Ads which were not served via either are removed as well as those living longer than 6 days to reduce noise from external traffic.

```{r data bucketing, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
input$ad_type <- ifelse(input$TOP.feature.ads==1 & input$DSS.feature.flag==1,"multi",
                       ifelse(input$TOP.feature.ads==1 & input$DSS.feature.flag==0,"TOP",
                              ifelse(input$TOP.feature.ads==0 & input$DSS.feature.flag==1,"DSP","none")))
input_no_none <- subset(input, ad_type!="none")
input <- subset(input_no_none, ad_days<7)
```

### Dealers' ads distribution Chi-square test
The representation of dealers between DSP and TOP ads is checked: the 2 distributions correlate with strengh approx 20%

```{r dealer distribution chi-test, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
dealers <- input[,c("dealer_id","ad_type")]
dealers <- melt(dealers)
dealers_cast <- dcast(dealers, dealer_id ~ ad_type)
dealers_cast <- dealers_cast[,-1]
#Chi-square test
chisq.test(dealers_cast)
#Cramér's V
assocstats(table(dealers_cast$DSP, dealers_cast$TOP))
```

### KPI calculation
TOP adVIPs are adjusted to 30% of their actual value to compensate for the extra traffic such placement received

```{r KPI calculation, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
input$TOPadVIPadj <- input$vipViews_TopFeature * 0.3
input$adVIP <- (input$TOPadVIPadj+input$vipViews_DSPtestFeature)
input$ad_days_DSP_TOP <- (input$TopFeature_days+input$DSPtestFeature_days)
input$adVIP_day <- input$adVIP/input$ad_days_DSP_TOP
input$totVIP_day <- input$vipViews_siteGermany/input$ad_days
input$adVIP_ad <- input$adVIP/(input$TOP.feature.ads+input$DSS.feature.flag)
input$totVIP_ad <- input$vipViews_siteGermany/(input$TOP.feature.ads+input$DSS.feature.flag)
```

### Outliers detection, exclusion
Outliers are calculated as data point exceeding the 999th permille on at least one for the main KPIs and excluded

```{r KPI outliers, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
outlier_threshold = 0.999
outlier_adVIP_day <- subset(input, adVIP_day>quantile(input$adVIP_day, outlier_threshold, na.rm = TRUE))
outlier_totVIP_day <- subset(input, totVIP_day>quantile(input$totVIP_day, outlier_threshold, na.rm = TRUE))
outlier_adVIP_ad <- subset(input, adVIP_ad>quantile(input$adVIP_ad, outlier_threshold, na.rm = TRUE))
outlier_totVIP_ad <- subset(input, totVIP_ad>quantile(input$totVIP_ad, outlier_threshold, na.rm = TRUE))
outlier_adVIP <- subset(input, adVIP>quantile(input$adVIP_day, outlier_threshold, na.rm = TRUE))
outlier_totVIP <- subset(input, TOPadVIPadj>quantile(input$totVIP_day, outlier_threshold, na.rm = TRUE))
outliers <- rbind(outlier_adVIP_day, outlier_totVIP_day, outlier_adVIP_ad, outlier_totVIP_ad, outlier_adVIP, outlier_totVIP)
input_clean <- anti_join(input, outliers, by=c("ad_id","dealer_id"))
```

### F-test in preparation for possible Student t-test
F-test is run to test for variance equality, a condition of the Student t-test, which would help comparing KPI distributions.
As the null hypothesis is rejected, variances aren't equal and the t-test cannot be performed

```{r F-test, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
input_clean_TOP <-  subset(input_clean, ad_type=="TOP") 
input_clean_TOP <- input_clean_TOP[,c("vipViews_siteGermany", "adVIP", "adVIP_day", "totVIP_day", "adVIP_ad", "totVIP_ad")]
input_clean_DSP <-  subset(input_clean, ad_type=="DSP") 
input_clean_DSP <- input_clean_DSP[,c("vipViews_siteGermany", "adVIP", "adVIP_day", "totVIP_day", "adVIP_ad", "totVIP_ad")]
#F-test 1
var.test(input_clean_TOP$vipViews_siteGermany,input_clean_DSP$vipViews_siteGermany)
#F-test 2
var.test(input_clean_TOP$adVIP,input_clean_DSP$adVIP)
```

### KPI average comparison
These values are to be taken as preliminary and only operational

```{r KPI avg comparison, echo=TRUE, warning=FALSE, tidy=TRUE, message=FALSE}
input_compare <- input_clean[,c("ad_type", "adVIP_day", "totVIP_day", "adVIP_ad", "totVIP_ad")] %>%
  group_by(ad_type) %>%
  summarise_all(mean)
```
```{r KPI avg table, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
library(knitr)
kable(input_compare, caption="Result")
```

## Plots:

### adVIP ~ totVIP (chart 1)

```{r chart 1, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean), aes(x=adVIP, y=vipViews_siteGermany, color=ad_type))+
  geom_point()+
  stat_smooth(method='lm', formula=y~x)
```

### Days displayed ~ totVIP (chart 2)

```{r chart 2, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean), aes(x=ad_days_DSP_TOP, y=vipViews_siteGermany, color=ad_type))+
  geom_point()+
  stat_smooth(method='lm', formula=y~x)
```

### Days displayed ~ totVIP - DSP (chart 3)

```{r chart 3, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean, ad_type=="DSP"), aes(x=ad_days_DSP_TOP, y=vipViews_siteGermany, group = cut_width(ad_days_DSP_TOP, 1)))+
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4)+
  coord_cartesian(ylim = c(0, 1000))
```

### Days displayed ~ totVIP - TOP (chart 4)

```{r chart 4, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean, ad_type=="TOP"), aes(x=ad_days_DSP_TOP, y=vipViews_siteGermany, group = cut_width(ad_days_DSP_TOP, 1)))+
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4)+
  coord_cartesian(ylim = c(0, 1000))
```

### Lightbox views ~ totVIP (chart 5)

```{r chart 5, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean, ad_type=="DSP"), aes(x=DSP.lightbox.views, y=vipViews_siteGermany, color=ad_type))+
  geom_point()+
  stat_smooth(method='lm', formula=y~x)+
  facet_wrap(~ad_days, scales='free')
```

### Lightbox views ~ totVIP by ad days (chart 6)

```{r chart 6, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean, ad_type!="multi"), aes(x=DSPdelivery, y=adVIP, color=ad_type))+
  geom_point()+
  stat_smooth(method='lm', formula=y~x)+
  facet_wrap(~ad_days, scales='free')
```

### adVIP ~ totVIP by ad days (chart 7)

```{r chart 7, echo=FALSE, warning=FALSE, tidy=TRUE, message=FALSE}
ggplot(subset(input_clean, ad_type!="multi"), aes(x=adVIP, y=vipViews_siteGermany, color=ad_type))+
  geom_point()+
  stat_smooth(method='lm', formula=y~x)+
  facet_wrap(~ad_days_DSP_TOP, scales='free')
```
